---
title: "Taheri_Cameron_HW02"
output: pdf_document
date: "2026-01-21"
---

```{r setup, message=FALSE, echo=FALSE}
#import packages
library(tidyverse)
```

### Population Means for Each Distribution

Normal Distribution $N(0,1)$ has mean = $\mu=0$ 

Chi-Squared Distribution $\chi^2(k),\ k=3$ has mean = $k=3$ 

Cauchy Distribution $C(0,1)$ does not have finite moments, but does have a median and mode, both equal to 0. 

### Generating Means
Generating sample means from increasingly large samples.

#### Function for generating the sample mean for each of the above distributions
```{r message=FALSE, echo=TRUE}
# calculate sample mean for each distribution
# input variable = i, parameter for distribution type
sample_mean <- function(i, dist_type = "norm") {
  n <- 2^i  # sample size
  
  # generate random samples from each distribution
  if (dist_type == "norm") {
    sample_data <- rnorm(
      n, 
      mean = 0, 
      sd = 1
      )
  } else if (dist_type == "chisqr") {
    sample_data <- rchisq(
      n, 
      df = 3
      )
  } else if (dist_type == "cauchy") {
    sample_data <- rcauchy(
      n, 
      location = 0, 
      scale = 1
      )
  }
  
  return(mean(sample_data))
}
```


#### Testing the sample mean function 
```{r message=FALSE, echo=FALSE}
N <- 20
dist_types <- c("norm", "chisqr", "cauchy")

# outer loop goes through each distribution
for (d in dist_types) {
  
  # inner loop generates mean using the sample_mean() function
  Ti <- numeric(N) 
  for (i in 1:N) {
    Ti[i] <- sample_mean(i, dist_type = d)
  }
  
  # create a temporary data frame for each distribution
  plot_df <- data.frame(
    i = 1:N,
    mean_val = Ti
  )
  
  # generate a plot and print
  p <- ggplot(
    data = plot_df, 
    aes(
      x = i, 
      y = mean_val
      )
    ) +
    
    geom_line(color = "steelblue") +
    geom_point(alpha = 0.5, color = "darkorange") +
    theme_minimal() +
    
    labs(
      title = paste("Sample Mean of:", d),
      x = "i (Sample Size = 2^i)",
      y = "Sample Mean (Ti)"
    )
  
  print(p)
}
```

We can see that for both the Normal and Chi-Squared distributions, the sample mean exists and converges to the population means of 0 and 3, respectively. However, for the Cauchy distribution, the sample mean is generated but it never converges. This is expected given the Cauchy does not have a finite first moment. 


### Simulations
Rerun the same distribution test 100 times for analyzing its behavior.

```{r message=FALSE, echo=FALSE}
N <- 20
S <- 100
dist_types <- c("norm", "chisqr", "cauchy")

for (d in dist_types) {
  
  # create empty plot for the lines
  # they're going to vary in size 
  #   given the characteristics of their distributions
  y_lims <- if(d == "cauchy") c(-50, 50) else if(d == "chisqr") c(0, 6) else c(-2, 2)
  
  plot(NULL, 
       xlim = c(1, N), 
       ylim = y_lims,
       xlab = "i (Sample Size = 2^i)", 
       ylab = "Sample Mean (Ti)",
       main = paste("100 Simulation Paths:", d))
  
  # outer loop will repeat nested loops 100 times
  for (s in 1:S) {
    
    # first inner loop grabs the sample mean for each distribution
    Ti_values <- numeric(N)
    for (i in 1:N) {
      Ti_values[i] <- sample_mean(i, dist_type = d)
    }
    
    # add the line to the empty plot
    lines(
      1:N, 
      Ti_values, 
      col = rgb(0, 0, 1, 0.2), 
      lwd = 0.5
      )
  }
  
}

```

We can see how this demonstrates the law of large numbers, whereby the sample mean will converge to the population mean should a population mean exist and be finite. In the case of the Cauchy distribution, the 100 simulations led to a cluster around zero (its median and mode) but with many observations far from zero. For the Normal and Chi-Squared, however, as the sample size increases (exponentially), the distribution converges around its population mean because each of them have a mean that exists and is finite.

One thing to note: I will be returning to this to find a better way to execute 100 simulations for each distribution (e.g., vectorizing).






